\paperTitle

### 2014-02-01 doc/icml2014/directed.tex
Many fine changes

We now show this condition is sufficient to learn the marginal
  distribution of the clique.
Consider any hidden clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$ for which this property holds. Let
  $x_{v_j}$ be an exclusive view for $h_{i_j}$ in $\sC$ (not necessarily unique) and $\sV
  = \{x_{v_1}, \cdots, x_{v_m}\}$ be a set of views for the clique $\sC$.
We can write down the moments $\mO_\sV$ as follows:
\begin{align*}
  \mO_\sV 
  &\eqdef \Pr(\Sx{\sV}) \\
  %&= \sum_{\vh \in \sH}
      %\Pr(\Sh{\sC}) \Pr(\Sx{\sV} \given \Sh{\sC}) \\
      &= \sum_{\vh \in \sH} \Pr(\Sh{\sC}) 
          \Pr(x_{i_1} | h_{v_1}) \cdots \Pr(x_{i_m} | h_{v_m}) \\
    &= Z_{\sC}(\mOpp{v_1}{i_1},\cdots,\mOpp{v_m}{i_m}).
\end{align*}

@@

  Let $x_v$ be an observed variable with parents $P$ and children $C$.
  Consider the following transformation.
  Replace $x_v$ with a new hidden variable \hp\ with the same
  parents $\Pa(x_v)$ and children $\text{Ch}(x_v) \union \{x_{v_1}, x_{v_2}, x_{v_3}\}$,
  where $x_{v_1},x_{v_2},x_{v_3}$ are three copies of $x$
  (\figureref{reduction}). 
  The domain of $\hp$ is the same as $x$,
    and $\mOpp{v}{\textrm{new}} = I$.\footnote{
      Though we have assumed that all hidden variables share the
      same domain, $[k]$, our work generalizes to hidden domains of
      different sizes, provided \assumptionref{full-rank-plus} is satisfied.
      }

@@

  Suppose $h$ had no unique view, then $\sV_{h} \subseteq \Union_{h' \in \sC^-} \sV_{h'}$. 
  However, this would imply that for every $x_1, x_2 \in \sV_{h}$, there
    exists some $h_1, h_2 \in \sC^-$ (not necessarily distinct) such that
    $x_1$ is connected to $h_1$ and $x_2$ is connected $h_2$.
  By participation in the clique, either $h_1 = h_2$ or $h_1$ is
    connected to $h_2$, implying that $x_1$ and $x_2$ are connected
    via a path that does not pass through $h$, contradicting the
    statement that $x_1, x_2 \in \sV$.

  Thus it must be the case that $\sV_{h} \subsetneq \Union_{h' \neq
    h \in \sC} \sV_{h'}$ and hence there exists a unique view for $h$.

  Algorithmically, finding the unique views can be done by subtracting
    out the remaining set, i.e. $x_{v_h} \in \sV_h \setminus \Union_{h'
    \in \sC} \sV_{h'}$
\end{proof}

With this lemma in place, we present the full algorithm, \LearnMarginals~
in \algorithmref{directed}.

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{algorithm}
  \caption{\LearnMarginals}
  \label{algo:directed}
  \begin{algorithmic}
    \REQUIRE A graphical model $\sG$ satisfying \propertyref{bottleneck}, data $\sD$
    \ENSURE Marginals $Z_\sC$ for every clique $\sC \in \sG$

@@

In the HMM (\figureref{examples-hmm}), the parameters
are $O \eqdef \Pr(x_i|h_i)$  and that $T \eqdef \Pr(h_{i+1} | h_i)$
for all $i$. % (i.e. we have parameter sharing).
While the first and last hidden variables $h_1, h_T$ in the
  sequence are not bottlenecks, they still have exclusive views ($x_1$ and
  $x_T$ respectively) whose parameters we know because they share
  parameters, $O$.
%We first use the bottleneck $h_2$ with views $x_1, x_2, x_3$ to estimate $O$,
%and then recover $\pi$ from $\{h_1\}$ and recover $T$ from the clique $\{h_{1}, h_{2}\}$.

@@

The simplest example of a model that cannot be recovered by our algorithm is
  a mixture model with a single view, i.e. \figureref{three-view} with
  just the variables $h_1$ and $x_1$.
However, without further assumptions the model itself is not
  identifiable\footnote{To see this, observe that the probability-mass
  can be arbitrarily exchanged between $\Pr(h_1)$ and $\Pr(x_1 | h_1)$}.
A class of binary-valued noisy-or networks
(see \figureref{examples-noisy-or} for an example)
which do not possess the
  bottleneck property are nonetheless identifiable;
  see \citet{halpern13noisyor} for an algorithm.

@@

From \citet{anandkumar13tensor} given \assumptionref{full-rank-plus},
we have that $\mOpphat{v}{i}$ converges to $\mOpp{v}{i}$ at a rate of $n^{-\frac12}$ with a constant that
depends polynomially on the $k$-th singular value of $\mOpp{v}{i}$.
%Of course, the singular value of $\mOpp{v}{i}$ become exponentially worse as $x_v$ and $h_i$ become farther.
%So while we can obtain consistent estimates, the sample complexity
%is exponential in the mimimum distance from $h_i$ to any three views, as to be expected.
%(\todo{say something
%quantitative?  annoying part is that we would have to bring in the bounds, so leave it out probably}).
% DONE: need to setup the asymptotic - it's not fresh in everyone's mind.
% Easy case
Next, we need to show that the hidden marginals $\hat Z_\sC$ converge to $Z_\sC$.
First, $\hat M_\sV$ is just an empirical average of multinomials over the data points.
Abusing notation slightly, we let $M_\sV$ also denote its $d^{|\sV|}$-dimensional vectorized version.
We also represent $\mH_\sC$ as
  a vector in $\Re^{k^m}$, and represent $\mOppAll \eqdef
  \mOpp{v_1}{i_1} \otimes \cdots \otimes
  \mOpp{v_m}{i_m}$ as a matrix in $\Re^{d^m \times
  k^m}$.
By the central limit theorem, we have:
$\sqrt{n} (\hat M_\sV - M_\sV) \convind \sN(0, \Sigma_\sV)$,
where $\Sigma_\sV$ is the \emph{asymptotic variance} of $\hat M_\sV$: 
% ARUN: I don't think so - \todo{doesn't this need to be multiplied by something?}
\begin{align*}
\Sigma_\sV \eqdef \dM_\sV - M_\sV M_\sV^\top, \quad \dM_\sV \eqdef \text{diag}(M_\sV).
\end{align*}

>>>>

We now show this condition is sufficient to learn the marginal
  distribution of the clique.
Consider any hidden clique $\sC = \{h_{i_1}, \cdots, h_{i_m}\} \in \sG$ for which this property holds. Let
  $x_{v_j}$ be any exclusive view for $h_{i_j}$ in $\sC$ and $\sV
  = \{x_{v_1}, \cdots, x_{v_m}\}$ be a set of views for the clique $\sC$.
We can write down the moments $\mO_\sV$ as follows:
\begin{align*}
  \mO_\sV 
  &\eqdef \Pr(\Sx{\sV}) \\
  %&= \sum_{\vh \in \sH}
      %\Pr(\Sh{\sC}) \Pr(\Sx{\sV} \given \Sh{\sC}) \\
      &= \sum_{\vh \in \sH} \Pr(\Sh{\sC}) 
          \Pr(x_{i_1} | h_{v_1}) \cdots \Pr(x_{i_m} | h_{v_m}) \\
    &= Z_{\sC}(\mOpp{v_1}{i_1},\cdots,\mOpp{v_m}{i_m}).
\end{align*}

@@

  Let $x_v$ be an observed variable with parents $\Pa(x_v)$ and children $\text{Ch}(x_v)$.
  Consider the following transformation.
  Replace $x_v$ with a new hidden variable \hp\ with the same
  parents $\Pa(x_v)$ and children $\text{Ch}(x_v) \union \{x_{v_1}, x_{v_2}, x_{v_3}\}$,
  where $x_{v_1},x_{v_2},x_{v_3}$ are three copies of $x$
  (\figureref{reduction}). 
  The domain of $\hp$ is the same as $x$,
    and $\mOpp{v}{\textrm{new}} = I$.\footnote{
      Though we have assumed that all hidden variables share the
      same domain, $[k]$, our work generalizes to hidden domains of
      different sizes, provided \assumptionref{full-rank-plus} is satisfied.
      }

@@

  Suppose $h$ had no unique view, then $\sV_{h} \subseteq \Union_{h' \in \sC^-} \sV_{h'}$. 
  However, this would imply that for every $x_1, x_2 \in \sV_{h}$, there
    exists some $h_1, h_2 \in \sC^-$ (not necessarily distinct) such that
    $x_1$ is connected to $h_1$ and $x_2$ is connected $h_2$.
  By participation in the clique, either $h_1 = h_2$ or $h_1$ is
    connected to $h_2$, implying that $x_1$ and $x_2$ are connected
    via a path that does not pass through $h$, contradicting the
    statement that $x_1, x_2 \in \sV_h$.

  Thus, it must be the case that $\sV_{h} \subsetneq \Union_{h' \neq
    h \in \sC} \sV_{h'}$ and hence there exists a unique view for $h$.

  Algorithmically, finding the unique view for $h$ can be done by
  subtracting the views of all other variables in the clique from
  $\sV_h$,i.e. $x_{v_h} \in \sV_h \setminus \Union_{h' \in \sC} \sV_{h'}$
\end{proof}

With this lemma in place, we present the full algorithm, $\LearnMarginals$,
in \algorithmref{directed}.

\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\begin{algorithm}
  \caption{\LearnMarginals}
  \label{algo:directed}
  \begin{algorithmic}
    \REQUIRE Graphical model $\sG$ satisfying \propertyref{bottleneck}, data $\sD$
    \ENSURE Marginals $Z_\sC$ for every clique $\sC \in \sG$

@@

In the HMM (\figureref{examples-hmm}), the parameters
are $O \eqdef \Pr(x_i|h_i)$  and that $T \eqdef \Pr(h_{i+1} | h_i)$
for all $i$. % (i.e. we have parameter sharing).
While the first and last hidden variables $h_1, h_M$ in the
  sequence are not bottlenecks, they still have exclusive views ($x_1$ and
  $x_M$ respectively) whose parameters we know because they share
  parameters, $O$.
%We first use the bottleneck $h_2$ with views $x_1, x_2, x_3$ to estimate $O$,
%and then recover $\pi$ from $\{h_1\}$ and recover $T$ from the clique $\{h_{1}, h_{2}\}$.

@@

The simplest example of a model that cannot be recovered by our algorithm is
  a mixture model with a single view, i.e. \figureref{three-view} with
  just the variables $h_1$ and $x_1$.
However, without further assumptions the model itself is not
  identifiable\footnote{To see this, observe that probability-mass
  can be arbitrarily exchanged between $\Pr(h_1)$ and $\Pr(x_1 | h_1)$ while preserving the marginals $\Pr(x_1)$}.
A class of binary-valued noisy-or networks
(see \figureref{examples-noisy-or} for an example)
which do not possess the
  bottleneck property are nonetheless identifiable;
  see \citet{halpern13noisyor} for an algorithm.

@@

From \citet{anandkumar13tensor} given \assumptionref{full-rank-plus},
we have that $\mOpphat{v}{i}$ converges to $\mOpp{v}{i}$ at a rate of $n^{-\frac12}$ with a constant that
depends polynomially on the $k$-th singular value of $\mOpp{v}{i}$.
%Of course, the singular value of $\mOpp{v}{i}$ become exponentially worse as $x_v$ and $h_i$ become farther.
%So while we can obtain consistent estimates, the sample complexity
%is exponential in the mimimum distance from $h_i$ to any three views, as to be expected.
%(\todo{say something
%quantitative?  annoying part is that we would have to bring in the bounds, so leave it out probably}).
% DONE: need to setup the asymptotic - it's not fresh in everyone's mind.
% Easy case
Next, we need to show that the hidden marginals $\hat Z_\sC$ converge to $Z_\sC$.
First, $\hat M_\sV$ is just an empirical average of multinomials over the data points.
Abusing notation slightly, we let $M_\sV$ also denote its $d^{m}$-dimensional vectorized version.
We also represent $\mH_\sC$ as
  a vector in $\Re^{k^m}$, and represent $\mOppAll \eqdef
  \mOpp{v_1}{i_1} \otimes \cdots \otimes
  \mOpp{v_m}{i_m}$ as a matrix in $\Re^{d^m \times
  k^m}$.
By the central limit theorem, we have:
$\sqrt{n} (\hat M_\sV - M_\sV) \convind \sN(0, \Sigma_\sV)$,
where $\Sigma_\sV$ is the \emph{asymptotic variance} of $\hat M_\sV$: 
% ARUN: I don't think so - \todo{doesn't this need to be multiplied by something?}
\begin{align*}
\Sigma_\sV \eqdef \dM_\sV - M_\sV M_\sV^\top, \quad \dM_\sV \eqdef \text{diag}(M_\sV).
\end{align*}

### 2014-02-01 doc/icml2014/intro.tex
Many fine changes

% 3. People have approached unsupervised learning with the MoM magic sauce, but the sauce is limited.
Recently, unsupervised learning techniques based on the spectral method of moments
have offered a refreshing perspective on
this learning problem
\citep{mossel2005learning,
hsu09spectral,
bailly2010spectral,
song2011spectral,
anandkumar11tree,anandkumar12moments,anandkumar12lda,
hsu12identifiability,
balle12automata}.
%\todo{make sure we are citing the right people}
These methods exploit the linear algebraic properties of the model to
factorize the moments into parameters, providing strong theoretical guarantees.
However, they apply to a limited set of models, and are thus
not as broadly applicable as EM.

@@

% 5. We get moments from third-order tensors from bottlenecks and factorize them into marginals,
% then get parameters.
% DONE: try to talk about the approach in terms of three steps
Our approach is illustrated in \figureref{approach}. 
First, we identify the views for each hidden variable $h_i$ (for example,
$x_1^a$, $x_1^b$ and $x_3^a$ are conditionally independent given $h_1$) and use
the tensor factorization algorithm of
\citet{anandkumar12moments,anandkumar13tensor} to identify the \emph{conditional
moments} $\E[x_i^a \mid h_i]$ and $\E[x_i^b \mid h_i]$ for each $i$.
Second, we optimize a \emph{composite marginal likelihood} to recover the marginals over
each hidden clique (e.g., $\mathbb P(h_2, h_3, h_4)$).
Normally, such a marginal likelihood objective would be non-convex,
but given the conditional moments, we obtain a convex objective,
which can be globally optimized using EM.
Finally, the hidden marginals yield the model parameters
via simple post-processing in the directed case and convex optimization in the undirected case.
%Given these factors, we show that there exists a set of observed
  %variables for every clique such that the likelihood is convex in the
  %parameters of that clique. 
%This guarantees that we can recover the true marginal distributions
  %using EM.
  %finally, with some post-processing, we are able to
  %consistently recover the true parameters of the model.
In \sectionref{directed}, we detail our algorithm and
show that our algorithm yields consistent parameter estimates
under non-degenerate rank assumptions on the conditional moments.
%describe the
  %settings and technical assumptions in which this approach is guaranteed
  %to work.
% PL: hard to talk about $O$ and $T$ at this stage.
%As a particular example, we are able to recover parameters for the
  %model depicted in \figureref{approach} if $O$ and $T$ are full-rank.

>>>>

% 3. People have approached unsupervised learning with the MoM magic sauce, but the sauce is limited.
Recently, unsupervised learning techniques based on the spectral method of moments
have offered a refreshing perspective on
this learning problem
\citep{mossel2005learning,
hsu09spectral,
bailly2010spectral,
song2011spectral,
anandkumar11tree,
anandkumar12lda,
anandkumar12moments,
hsu12identifiability,
balle12automata}.
%\todo{make sure we are citing the right people}
These methods exploit the linear algebraic properties of the model to
factorize the moments into parameters, providing strong theoretical guarantees.
However, they apply to a limited set of models, and are thus
not as broadly applicable as EM.

@@

% 5. We get moments from third-order tensors from bottlenecks and factorize them into marginals,
% then get parameters.
% DONE: try to talk about the approach in terms of three steps
Our approach is illustrated in \figureref{approach}. 
First, we identify the views for each hidden variable $h_i$ (for example,
$x_1^a$, $x_1^b$ and $x_3^a$ are conditionally independent given $h_1$) and use
the tensor factorization algorithm of
\citet{anandkumar13tensor} to identify the \emph{conditional
moments} $\Pr[x_i^a \mid h_i]$ and $\Pr[x_i^b \mid h_i]$ for each $i$.
Second, we optimize a \emph{composite marginal likelihood} to recover the marginals over
each hidden clique (e.g., $\mathbb P(h_2, h_3, h_4)$).
Normally, such a marginal likelihood objective would be non-convex,
but given the conditional moments, we obtain a convex objective,
which can be globally optimized using EM.
Finally, the hidden marginals yield the model parameters
via simple post-processing in the directed case and convex optimization in the undirected case.
%Given these factors, we show that there exists a set of observed
  %variables for every clique such that the likelihood is convex in the
  %parameters of that clique. 
%This guarantees that we can recover the true marginal distributions
  %using EM.
  %finally, with some post-processing, we are able to
  %consistently recover the true parameters of the model.
In \sectionref{directed}, we detail our algorithm and
show that our algorithm yields consistent parameter estimates
under non-degenerate rank assumptions on the conditional moments.
%describe the
  %settings and technical assumptions in which this approach is guaranteed
  %to work.
% PL: hard to talk about $O$ and $T$ at this stage.
%As a particular example, we are able to recover parameters for the
  %model depicted in \figureref{approach} if $O$ and $T$ are full-rank.
